Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management

### 2021-07-26 Scripts/Supplements.Rmd
Sent to co-authors

**Materials and methods**

_**Databases and species selection**_

@@

_**Projecting species distributions under climate change**_

@@

_**Calculating an index of transboundary range shift**_

@@

_**Calculating shifts in the shared distribution of transboundary stocks**_

@@

_**Quantifying the intensity of transboundary stocks’ range shift**_

@@

First, we estimated **the catch proportion as the** stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100. We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv]. Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$). These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN]. The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$. Thus, a change of SSR catch beyond an EEZ’s threat point happens when the future SSR catch exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];

@@

_**Statistical analysis**_

We tested the results for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz]  to investigate geopolitical and ecological differences in the range shift of transboundary species. Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift. In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs. All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm]. Code for all analyses is available at https://github.com/jepa/TransEmergence

@@

_**Figures**_

>>>>

# Materials and methods

### Databases and species selection

@@

### Projecting species distributions under climate change

@@

### Calculating an index of transboundary range shift

@@

### Calculating shifts in the shared distribution of transboundary stocks

@@

### Quantifying the intensity of transboundary stocks’ range shift

@@

First, we estimated the catch proportion as the stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100. We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv]. Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$). These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN]. The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$. Thus, a change of SSR catch beyond an EEZ’s threat point happens when the future SSR catch exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];

@@

### Statistical analysis

We tested the results for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz]  to investigate geopolitical and ecological differences in the range shift of transboundary species. Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift. In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs. All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm]. Code for all analyses is available at https://github.com/jepa/EmergingFish

@@

# Supplemental Figures
